Optimal (Multiway) Spatial Joins

### 2024-05-20 msj.tex
yufei

Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary. The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem. Specifically, if the 3D version of the 3-SJ problem could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}. This can be shown using an argument in \cite{kcko22}, which, however, was written to prove a more generic result. We simply the argument for 3-SJ and present it in full in Appendix~\ref{app:lb-cond}.

>>>>

Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary. The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem. Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}. This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result. We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.
